The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2016-12-07 DishBeamMeasurements/paper/DishBeamMeasurementsPaper_rev5.tex
update to published version

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA) is a radio interferometer aiming to detect the power spectrum of 21\,cm fluctuations from neutral hydrogen from the Epoch of Reionization (EOR). Drawing on lessons from the Murchison Widefield Array (MWA) and the Precision Array for Probing the Epoch of Reionization (PAPER), HERA is a hexagonal array of large (14\,m diameter) dishes with suspended dipole feeds. Not only does the dish determine overall sensitivity, it affects the observed frequency structure of foregrounds in the interferometer. This is the first of a series of four papers characterizing the frequency and angular response of the dish with simulations and measurements. We focus in this paper on the angular response (i.e., power pattern), which sets the relative weighting between sky regions of high and low delay, and thus, apparent source frequency structure. We measure the angular response at 137\,MHz using the ORBCOMM beam mapping system of \citet{neben15}. We measure a collecting area of 93\,m$^2$ in the optimal dish/feed configuration, implying HERA-320 should detect the EOR power spectrum at $z\sim9$ with a signal-to-noise ratio of 19.3 using a foreground avoidance approach, and 74.3 using a foreground subtraction approach. Lastly we study the impact of these beam measurements on the distribution of foregrounds in Fourier space.
\end{abstract}

@@

The HERA element (Fig. \ref{fig:feeddiagram}) is a 14\,m diameter faceted parabolic dish ($ f/D=0.32$) with a dual-polarized dipole feed suspended at prime focus \citep{dishmemo}. Here $f$ is the focal length of the dish and $D$ is the dish diameter. The dish surface is formed by wire mesh sheets (i.e., facets) mounted on PVC tubes which run from the lip of the dish to the hub at the vertex. For these tests, the feed consists of a dual linear polarization PAPER sleeved dipole mounted 17'' below a 78'' diameter wire mesh back plane surrounded by a 30'' deep cylinder, suspended from cables attached to each of the three telephone poles around the dish. The dipole ``sleeves'' are circular disks just above and below the dipole designed to broaden its frequency response. The cylinder is offset 0.5'' from the back plane, and is designed to make the beams for both linear polarizations more similar. Fig. \ref{fig:feedphoto} shows the feed as deployed on the ground for early testing.  The nominal dish focus is $f=(f/d)D=4.48$\,m, though given its faceted design, the dish does not have a single focus. Our numerical electromagnetic models suggest the best focus is slightly higher due to the feed geometry. In this work we study the dish beam pattern at rigging heights of 4.5\,m, 5.0\,m, and 5.3\,m, measured from dish surface to feed plane, the last height being the maximum height we can achieve with the feed suspension system installed on the dish. These height measurements are uncertain at the $\pm5\%$ level in this study. For more details on the dish design and construction see \citet{deboer16}. Feed/dish optimization studies are ongoing and the values of these parameters may change in the full HERA array \citep{feedoptimizationmemo}.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.jpg}
\caption{Prototype HERA feed seen here outside the dish and upside-down for preliminary characterization. This feed revision consists of a dual-polarized sleeved dipole offset 17'' from a 78'' diameter back plane, surrounded by a 30'' deep cylindrical skirt.}
\label{fig:feedphoto}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=7in]{dave195_rich195_airy_beams.pdf}
\caption{Simulated dish power patterns (NS polarization) at 137\,MHz (see Sec. \ref{sec:dishmodels}) with $h_\text{feed}=5$\,m using HFSS (left) and CST (middle) are shown beside an ideal Airy pattern for a 14\,m diameter dish for comparison. Dased lines mark zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, and 80$^\circ$.}
\label{fig:modelbeams}
\end{figure*}

We numerically model the HERA dish in both ANSYS HFSS\footnote{http://www.ansys.com/Products/Electronics/ANSYS-HFSS} and in CST Microwave Studio\footnote{https://www.cst.com/Products/CSTMWS}, making slightly different assumptions, in order to study the range of realistic beams given modeling inaccuracies and material imperfections. In particular, the near horizon beam response, which sets the level of horizon brightening in the delay spectrum, is quite sensitive to modeling assumptions.

The HFSS model uses a realistic faceted dish \citep{deboer16} ($f/D=0.32$) with a 1\,m diameter dielectric similar to dry soil at the vertex. In reality a 1\,m diameter circle at the vertex is filled with concrete and supports the PVC pipes which suspend the mesh panels, but we effectively assume here it is perfectly absorbant.  The feed is modeled as a sleeved dipole with a solid metal back plane and cylinder with the dimensions given above. Copper is used for the dipole itself, and aluminum for all reflecting surfaces. In the simulations, we excite one of the dipoles using a modal port and measure the total gain response in each direction to 137.5\,MHz radiation.

The CST model assumes a 14\,m diameter perfect paraboloid of solid aluminum with $f/D=0.32$ but neglects the 1\,m hole at the vertex, effectively assuming the concrete and earth ground behind it are perfectly reflective. The feed model has the same dimensions and properties as in the HFSS model. These simulations are done by exciting one of the dipoles with band-limited noise (100--200\,MHz), then measuring the farfield radiation in each direction after the excitation pulse energy within the structure decays to below -80\,dB.

The simulated HFSS and CST beams for the NS dipole are plotted in Fig. \ref{fig:modelbeams} (left and center panel) along with an Airy pattern for comparison. As expected, both model beams have slightly stronger sidelobes and wider main lobes than the ideal Airy pattern. The dipole sleeve (circular pieces in Fig. \ref{fig:feedphoto}) and skirt result in a feed beam which is slightly elongated in the E plane and slightly compressed in the H plane, opposite to the behavior of a simple dipole. This wider dish illumination in the NS direction by the NS feed dipole results in a narrower dish beam in the NS direction. Similarly, the EW dish beam is narrower in the EW direction. Lastly, we note that in both models, the best focus is found to be close to 5.23\,m with this feed geometry.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{ref_dipole_and_hera_dish.jpg}
\caption{The dish with its suspended feed is seen in the back, 50\,m north of one of the reference antennas used in the null experiment to study systematics. The experiment is conducted in Galford Meadow at NRAO--Green Bank.}
\label{fig:greenbankdishphoto}
\end{figure}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.5in]{measured_beams_and_models_slices.pdf}
\caption{Slices through the E (red) and H (cyan) planes through the measured dish power patterns (points) and numerical models (curves). The measured beams agree with both models in the main lobe out to zenith angles of 15--20$^\circ$ up to slight main lobe tilts, but begin to deviate in the sidelobes where the beam response is 25-30\,dB down from zenith. The measured beams typically differ more from both model beams than the models differ from each other, suggesting that real world effects are more significant that the slightly different assumptions used in the CST and HFSS simulations. In particular, the most likely systematic is mis-centering of the feed over the dish (see Sec. \ref{sec:powerpatternmeasurements}).}
\label{fig:measuredbeamslices}
\end{figure*}

@@

Figure \ref{fig:measuredbeamslices} shows slices through the E and H planes of these power patterns along with the HFSS and CST numerical models discussed earlier. As in the previous plot, the EW and NS beams are shown in the left and right panels, while the different feed heights are shown in the different rows. The data agree with both models to within 1\,dB in the main lobe, though in several cases appear slightly shifted so they are not quite centered on zenith. The data diverge further in the sidelobes at zenith angles of $20^\circ$ and larger. Here the evolution of the sidelobes as the feed is raised is again seen starkly, as is the fact that the main lobes are slightly wider along the H planes than along the E planes. We observe that both models agree with the measured beams in the main lobe but deviate from the data in different ways at the 1--5 dB level in the sidelobes. Neither model looks consistently better, suggesting that the as-built HERA dish deviates from both numerical models more significantly than the models deviate from each other due to their slightly different assumptions.

We emphasize that the model deviations observed in the measured beams are real in that they are larger than the 0.5\,dB scale systematics observed in the null experiments (Fig. \ref{fig:nullexptplots}). Those experiments bound the impact of environmental reflections and reference dipole mismodeling to the 10\% level or smaller across the whole sky. The observed dish beam asymmetries, model deviations in sidelobes, and slight shifts of the main lobes all suggest feed centering errors. The feed is suspended by rope from three telephone poles spaced around the dish, and is raised by pulling all three ropes to a new length. Each time this is done the feed centering is slightly disturbed because all three ropes must be pulled to the exact same length to center the feed. If feed tilt/rotation errors or dish surface imperfections were significant, then the beam errors at different feed heights would look similar because these errors persist when the feed is raised or lowered. All three ropes are tied to a single location on the feed, so rope length errors affect on the feed centering, not its rotation or tilt. The fact that the observed model deviations change over the three beam measurements suggests that feed centering errors are most significant because they are the only ones which change when the feed height is changed. To mitigate these errors, the feeds in the full HERA array will be tied down to the dish surface at several points.

@@

The collecting areas are shown in Table 1 along with the maximal collecting area achieved by the Airy pattern for a 14\,m dish. The measured collecting areas are 30--50\% lower than the geometric area. This is in line with expectations given that the Airy pattern has the largest possible collecting area, equal to the dish cross section, and the feed's backplane and cylindrical skirt reduce it. However we opt for this reduction over added collecting area in order to reduces the azimuthal beam asymmetry.

 \begin{table}[h]
 \caption{ \label{table:collectingareatable}Collecting area (m$^2$) of measured 137\,MHz beams and corresponding power spectrum SNR for HERA-320 using either foreground avoidance or foreground subtraction.}
\begin{tabular}{| l | l | l |}
\hline
Beam & $A_\text{eff}$ (m$^2$) & SNR ($\sigma$)\\
&& (avoidance, subtraction)\\
\hline
  Airy pattern & 155 & 25.5, 90.8  \\
    Measured, feed at 5.3\,m & 93.0 & 19.3, 74.3 \\
    Measured, feed at 5\,m & 77.1 &  16.4, 67.9 \\
    Measured, feed at 4.5\,m & 68.5 &  15.4, 63.9 \\ 
  \hline
\end{tabular}
\end{table}

% To input these collecting areas into 21cmSense, we convert these measured dish collecting areas into effective dish diameters, which we input as the \texttt{dish\_size\_in\_lambda} parameter. 
We run 21cmSense\footnote{https://github.com/jpober/21cmSense} to compute the overall SNR of a power spectrum detection with one season (6 hours per night for 180 nights) of HERA-320 data. We use a fiducial Epoch of Reionization model generated with 21cmFast \citep{21cmfast}. This model assumes $\zeta=31.5$ for the ionizing efficiency, $T_\mathrm{vir}=1.5\times10^4$\,K for the minimum virial temperature of halos producing ionizing photons, and $R_\text{mfp}=30$\,Mpc for the mean free path of ionizing photons, and reaches 50\% ionization at $z \sim 9.5$ and complete ionization at $z \sim 7$, and is consistent with most current observations \citep[e.g.][]{PoberNextGen}. We predict SNRs first for a foreground \textit{avoidance} approach where only modes outside of the wedge plus a buffer of $\Delta k_\parallel=0.1\,\mathrm{h}\,\mathrm{Mpc}^{-1}$ are used. These modes have frequency dependence larger than that of any source on the sky. We also predict SNRs for a foreground \textit{subtraction} approach using all modes whose instrumental frequency dependence is larger than that of a source at the edge of the main lobe.

The SNRs computed with the measured collecting areas are 15-20 with foreground avoidance compared with 25 for the Airy pattern. With foreground subtraction, the SNR falls from 90 with the Airy pattern to 60-75 with the measured collecting areas. In all cases this reduction is a loss of sensitivity, but a power spectrum detection is still always very significant at the 15$\sigma$ level or better.

@@

We simulate foreground power spectra using different primary beam models at various local sidereal times (LSTs). We use frequency-independent model beams (evaluated at 137\,MHz) to isolate the interferometric foreground frequency dependence. The added frequency dependence of the changing overall gain and beam shape with frequency is addressed by the other papers in this series. Given that our measured dish power patterns agree well with both numerical models (HFSS and CST) in the main lobe but deviate in the sidelobes, and that these models make somewhat different assumptions about the dish surface, we take them as a representative pair of possible dish models. We use the empirically best feed height of 5.3\,m. We also include the Airy pattern for comparison as in \citet{nithya15}. Beam models with weaker response near the horizon (such as the Airy pattern) downweight sources in this direction of high apparent frequency dependence. This reduces the magnitude of emission near the edge of the EOR window, reducing the risk it leaks inside. We use the per-baseline approach of \citet{parsons12a,parsons12b} by first simulating visibilities measured by specific baselines as a function of frequency, then computing the Fourier transform over frequency (delay transform), and lastly normalizing the result into a cosmological power spectrum following \citet{nithya15}.

@@

Figure \ref{fig:delayspec} (top panel) shows simulated foreground delay spectra at various LSTs using the nominal HFSS beam. As all these LSTs correspond to high galactic latitudes far from the galactic center, the total visibility power (the level of the zero delay mode) varies only by a factor of a few over these LSTs on both baseline lengths (14.6\,m (left panel), 43.8\,m (right panel)). However the positive delay horizon limit (corresponding to the western horizon) has a peak that varies by over 1.5 orders of magnitude on both baselines, demonstrating the stark difference in horizon brightening when the galaxy is just above versus just below the horizon. In this figure we perform the approximate conversion from delay $\tau$ to $k_\parallel$ at $z=8$, which we plot as a second $x$-axis at the top of the plot.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{nithya_fg_pspec_all_lst.pdf}
\includegraphics[width=6in]{nithya_fg_pspec_lst0deg.pdf}
\includegraphics[width=6in]{nithya_fg_pspec_lst60deg.pdf}
\caption{We plot simulated foreground delay spectra using the HFSS beam at various LSTs (top panel). The maximum horizon brightening at the positive horizon occurs close to 0$^\circ$ LST. At this LST, the simulated foreground delay spectra for the three beam models differ markedly near the positive horizon, plotted as a vertical line at the baseline's maximum delay. In contrast, when the horizon brightening effect is smaller at 60$^\circ$ LST (bottom panel), the foreground delay spectra from all three beams agree better.}
\label{fig:delayspec}
\end{figure*}

@@

How much do the predicted foreground power spectra differ between the three model dish power patterns? Figure \ref{fig:delayspec} (middle panel) shows the simulated delay spectra for all three beams at $0^\circ$ LST, when the horizon brightening is worst. Both numerical models agree out to delays of roughly 20\,ns on the 14.6\,m baseline and 50\,ns on the 43.8\,m baseline. These numbers suggest that the beams track each other fairly well out to roughly 25$^\circ$ from zenith, beyond which they diverge. This is roughly what is observed in Figure \ref{fig:measuredbeamslices}. At larger delays, especially near the positive delay horizon limit, all three model delay spectra diverge due to the significant edge brightening which effectively discriminates between these models. The CST and HFSS beams reach roughly -32\,dB and -38\,dB at $80^\circ$ zenith angle (Figure \ref{fig:modelbeams}), consistent with the fact that the CST beam has a larger horizon brightening than the HFSS beam. This is seen in the delay spectra for both baseline lengths, though the edge brightening is much clearer on the longer baseline where it less diluted by zero delay emission.

@@

We have presented beam pattern measurements at 137\,MHz, and discussed their implications for 21\,cm power spectrum analyses in terms of sensitivity and foreground isolation. We begun with power pattern measurements made using the beam mapping system of \citet{neben15} which we deployed at the prototype three-element HERA array at the National Radio Astronomy Observatory--Green Bank. We measured the dish power pattern with the feed at different heights over the dish surface and found that the best focus is at a feed rigging height of 5.3\,m, though this may change for different feed designs being explored\citep{feedoptimizationmemo}. The measured beams probe nearly two thirds of the visible sky down to -30\,dB relative to boresight, and agree well with both models in the main lobe out to 10--20$^\circ$ from zenith. The measured beams roughly track the predicted sidelobe levels at 20--30\,dB below zenith, deviating at the 1--5\,dB level.

These deviations away from models and away from 180$^\circ$ azimuthal symmetry are larger than the $\pm0.5$\,dB systematics observed in our null experiments which probe the accuracy of our beam measurement system, suggesting they are genuine measurements of the in situ dish beam. The most likely culprit is feed mis-centering which shifts and distorts the main lobe sidelobes. In the full HERA array, the suspended feeds will be tied to the dish surface at several points to level and center them. Characterizing the level of antenna-to-antenna beam variation in the full HERA array and its effects on power spectrum analyses, as Neben et al. (in press) do for the MWA, is left as future work.

We quantify HERA's sensitivity to the 21\,cm power spectrum given our beam measurements by first computing the collecting area of the measured beams at the different feed rigging heights, finding 93\,m$^2$ at the best focus. We observe a reduction of 30--50\% from the geometric cross section of a 14\,m dish due to feed blockage and under-illumination of the dish. We convert our measured collecting areas into effective dish sizes, then use 21cmSense to predict the overall power spectrum SNR at $z\sim9.5$ with one season of HERA-320 data. We predict SNRs of 19.3 and 74.3 using foreground avoidance and subtraction approaches respectively, compared with SNRs of 25.5 and 90.8 using an ideal unobstructed 14\,m aperture (Airy pattern). Still, these sensitivities permit a very significant detection of the 21\,cm signal after a single observing season.

@@

We predict the magnitude of this effect for the HERA element as well as its dependence on beam modeling uncertainties. As expected, we find that the level of horizon brightening is largest when the galaxy is just above the horizon, and lowest when it is well below. When this pitchfork effect is large, we find that the uncertainty in its predicted amplitude is also large, as seen in the differences between the delay spectra calculated using HFSS and CST beam models. When the effect is small, the two beam models produce much more similar results, highlighting the delay spectrum as an exquisite probe of the difficult-to-measure beam response at very low elevations.

>>>>

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA) is a radio interferometer aiming to detect the power spectrum of 21\,cm fluctuations from neutral hydrogen from the Epoch of Reionization (EOR). Drawing on lessons from the Murchison Widefield Array (MWA) and the Precision Array for Probing the Epoch of Reionization (PAPER), HERA is a hexagonal array of large (14\,m diameter) dishes with suspended dipole feeds. Not only does the dish determine overall sensitivity, it affects the observed frequency structure of foregrounds in the interferometer. This is the first of a series of four papers characterizing the frequency and angular response of the dish with simulations and measurements. We focus in this paper on the angular response (i.e., power pattern), which sets the relative weighting between sky regions of high and low delay, and thus, apparent source frequency structure. We measure the angular response at 137\,MHz using the ORBCOMM beam mapping system of \citet{neben15}. We measure a collecting area of 93\,m$^2$ in the optimal dish/feed configuration, implying HERA-320 should detect the EOR power spectrum at $z\sim9$ with a signal-to-noise ratio of 12.7 using a foreground avoidance approach with a single season of observations, and 74.3 using a foreground subtraction approach. Lastly we study the impact of these beam measurements on the distribution of foregrounds in Fourier space.
\end{abstract}

@@

The HERA element (Fig. \ref{fig:feeddiagram}) is a 14\,m diameter faceted paraboloid ($ f/D=0.32$) with a dual-polarized dipole feed suspended at prime focus \citep{dishmemo}. Here $f$ is the focal length of the dish and $D$ is the dish diameter. The dish surface is formed by wire mesh sheets (i.e., facets) mounted on PVC tubes which run from the lip of the dish to the hub at the vertex. For these tests, the feed consists of a dual linear polarization PAPER sleeved dipole mounted 17'' below a 78'' diameter wire mesh back plane surrounded by a 30'' deep cylinder. The feed is suspended from a single point on its back plane from three ropes, each attached to a telephone pole. The three telephone poles are equally spaced around the dish. The dipole ``sleeves'' are circular disks just above and below the dipole designed to broaden its frequency response. The feed cylinder is offset 0.5'' from the back plane, and is designed to make the dipole beam more azimuthally symmetric and also taper its response near the edges of the dish to reduce spillover into adjacent dishes. Fig. \ref{fig:feedphoto} shows the feed as deployed on the ground for early testing.

The nominal dish focus is $f=(f/D)D=4.48$\,m, though given its faceted design, the dish does not have a single focus. Our numerical electromagnetic models suggest the best focus is slightly higher than that of an perfect paraboloid. In this work we study the dish beam pattern at rigging heights of 4.5\,m, 5.0\,m, and 5.3\,m, measured from dish surface to feed plane, the last height being the maximum height we can achieve with the feed suspension system installed on the dish. These height measurements are uncertain at the $\pm5\%$ level in this study. For more details on the dish design and construction see \citet{deboer16}. Feed optimization studies are ongoing and the values of these parameters may change in the full HERA array \citep{feedoptimizationmemo}.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.pdf}
\caption{Prototype HERA feed seen here outside the dish and upside-down for preliminary characterization. This feed revision consists of a dual-polarized sleeved dipole offset 17'' from a 78'' diameter back plane, surrounded by a 30'' deep cylindrical skirt.}
\label{fig:feedphoto}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=7in]{dave195_rich195_airy_beams.pdf}
\caption{Simulated dish power patterns (NS polarization) at 137\,MHz (see Sec. \ref{sec:dishmodels}) with $h_\text{feed}=5$\,m using the full-faceted model (left) and the perfect paraboloidal model (middle) are shown beside an ideal Airy pattern for a 14\,m diameter dish for comparison. Dased lines mark zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, and 80$^\circ$.}
\label{fig:modelbeams}
\end{figure*}

We numerically model the HERA dish in two different ways in order to study the range of realistic beams given modeling inaccuracies and material imperfections. In particular, the near horizon beam response, which sets the level of horizon brightening in the delay spectrum, is quite sensitive to modeling assumptions. We first generate a full-faceted model of the dish using ANSYS HFSS\footnote{http://www.ansys.com/Products/Electronics/ANSYS-HFSS}. All mesh surfaces are modeled as solid aluminum and the dipole itself is modeled as copper. The 1\,m concrete circle at the vertex is modeled with a dielectric similar to dry soil. For comparison, we also model the dish as a perfect paraboloid. We simulate this second model using CST Microwave Studio\footnote{https://www.cst.com/Products/CSTMWS}, but the differences are dominated by the dish geometry, not the choice of numerical electromagnetic solver.

The simulated full-faceted and perfect paraboloid beams for the NS dipole are plotted in Fig. \ref{fig:modelbeams} (left and center panel) along with an Airy pattern for comparison. As expected, both model beams have slightly stronger sidelobes and wider main lobes than the ideal Airy pattern. The dipole sleeve (circular pieces in Fig. \ref{fig:feedphoto}) and skirt result in a feed beam which is slightly elongated in the E plane and slightly compressed in the H plane, opposite to the behavior of a simple dipole. This wider dish illumination in the NS direction by the NS feed dipole results in a narrower dish beam in the NS direction. Similarly, the EW dish beam is narrower in the EW direction. Lastly, we note that in both models, the best focus is found to be close to 5.23\,m with this feed geometry.

@@

\begin{figure}[h]
\includegraphics[width=3.4in]{ref_dipole_and_hera_dish.pdf}
\caption{The dish with its suspended feed is seen in the back, 50\,m north of one of the reference antennas used in the null experiment to study systematics. The experiment is conducted in Galford Meadow at NRAO--Green Bank.}
\label{fig:greenbankdishphoto}
\end{figure}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.5in]{measured_beams_and_models_slices.pdf}
\caption{Slices through the E (red) and H (cyan) planes through the measured dish power patterns (points) and numerical models (curves). The measured beams agree with both models in the main lobe out to zenith angles of 15--20$^\circ$ up to slight main lobe tilts, but begin to deviate in the sidelobes where the beam response is 25-30\,dB down from zenith. The measured beams typically differ more from both model beams than the models differ from each other, suggesting that real world effects are more significant that the slightly different assumptions used the two beam models. In particular, the most likely systematic is mis-centering of the feed over the dish (see Sec. \ref{sec:powerpatternmeasurements}).}
\label{fig:measuredbeamslices}
\end{figure*}

@@

Figure \ref{fig:measuredbeamslices} shows slices through the E and H planes of these power patterns along with the full-faceted and perfect paraboloid numerical models discussed earlier. As in the previous plot, the EW and NS beams are shown in the left and right panels, while the different feed heights are shown in the different rows. The data agree with both models to within 1\,dB in the main lobe, though in several cases appear slightly shifted so they are not quite centered on zenith. The data diverge further in the sidelobes at zenith angles of $20^\circ$ and larger. Here the evolution of the sidelobes as the feed is raised is again seen starkly, as is the fact that the main lobes are slightly wider along the H planes than along the E planes. We observe that both models agree with the measured beams in the main lobe but deviate from the data in different ways at the 1--5 dB level in the sidelobes. Neither model agrees consistently better with the data, suggesting that real-world imperfections of the HERA dish dominate over the slightly different modeling assumptions.

We emphasize that the model deviations observed in the measured beams are real in that they are larger than the 0.5\,dB scale systematics observed in the null experiments (Fig. \ref{fig:nullexptplots}). Those experiments bound the impact of environmental reflections and reference dipole mismodeling to the 10\% level or smaller across the whole sky. The observed dish beam asymmetries, model deviations in sidelobes, and slight shifts of the main lobes all suggest feed centering errors. The feed is suspended by three ropes attached from the center of the feed back plane to three telephone poles spaced around the dish, and is raised by pulling all three ropes to a new length. Each time this is done the feed centering is slightly disturbed because all three ropes must be pulled to the exact same length to center the feed. Because all three ropes are attached to the same point on the feed, changing their lengths does not affect feed rotation or tilt. Thus if rotation or tilt errors, or dish surface imperfections, were significant, then the beam errors at different feed heights would look similar. The fact that the observed model deviations change with feed height suggests that feed centering errors are most significant. To mitigate all these feed positioning errors, the feeds in the full HERA array will be tied down to the dish surface at several points.

@@

The collecting areas are shown in Table 1 along with the maximal collecting area achieved by the Airy pattern for a 14\,m dish. The measured collecting areas imply aperture efficiencies of 45--60\%. This is in line with expectations given the feed design which tapers the dipole beam towards the edges of the dish to reduce spillover into adjacent dishes. The mesh cylinder hanging from the feed back plane around the dipole also reduces the aperture efficiency slightly in order to make the feed beam more azimuthally symmetric.

 \begin{table}[h]
 \caption{ \label{table:collectingareatable}Collecting area (m$^2$) of measured 137\,MHz beams and corresponding power spectrum SNR for HERA-320 using either foreground avoidance or foreground subtraction.}
\begin{tabular}{| l | l | l |}
\hline
Beam & $A_\text{eff}$ (m$^2$) & SNR ($\sigma$)\\
&& (avoidance, subtraction)\\
\hline
  Airy pattern & 155 & 18.7, 90.8  \\
    Measured, feed at 5.3\,m & 93.0 & 12.7, 74.3 \\
    Measured, feed at 5\,m & 77.1 &  10.6, 67.9 \\
    Measured, feed at 4.5\,m & 68.5 &  10.0, 63.9 \\ 
  \hline
\end{tabular}
\end{table}

% To input these collecting areas into 21cmSense, we convert these measured dish collecting areas into effective dish diameters, which we input as the \texttt{dish\_size\_in\_lambda} parameter. 
We run 21cmSense\footnote{https://github.com/jpober/21cmSense} to compute the overall SNR of a power spectrum detection with one season (6 hours per night for 180 nights) of HERA-320 data. We use a fiducial Epoch of Reionization model generated with 21cmFast \citep{21cmfast}. This model assumes $\zeta=31.5$ for the ionizing efficiency, $T_\mathrm{vir}=1.5\times10^4$\,K for the minimum virial temperature of halos producing ionizing photons, and $R_\text{mfp}=30$\,Mpc for the mean free path of ionizing photons, and reaches 50\% ionization at $z \sim 9.5$ and complete ionization at $z \sim 7$, and is consistent with current observations \citep[e.g.][]{PoberNextGen}.

We predict SNRs first for a foreground \textit{avoidance} approach where only modes outside of the wedge plus a buffer of $\Delta k_\parallel=0.15\,\mathrm{h}\,\mathrm{Mpc}^{-1}$ are used. These modes have frequency dependence larger than that of any smooth spectrum source on the sky, and this buffer size is chosen to exclude modes which leak out of the wedge due to beam frequency dependence. Due to imperfect impedance matching at the center of the 100-200\,MHz band, the $z\sim8.5$ band requires a slightly larger buffer, though our chosen buffer effectively avoids the leakage in other bands \citep{ewallwice16}. We also predict SNRs for a foreground \textit{subtraction} approach using all modes whose instrumental frequency dependence is larger than that of a source at the edge of the main lobe.

The SNRs computed with the measured collecting areas are 10-13 with foreground avoidance compared with 19 for the Airy pattern. With foreground subtraction, the SNR falls from 90 with the Airy pattern to 60-75 with the measured collecting areas. In all cases this reduction is a loss of sensitivity, but a power spectrum detection is still always very significant at the 10$\sigma$ level or better.

@@

We simulate foreground power spectra using different primary beam models at various local sidereal times (LSTs). We use frequency-independent model beams (evaluated at 137\,MHz) to isolate the interferometric foreground frequency dependence. The added frequency dependence of the changing overall gain and beam shape with frequency is addressed by the other papers in this series. Given that our measured dish power patterns agree well with both numerical models (full-faceted and perfect paraboloid) in the main lobe but deviate in the sidelobes, and that these models make somewhat different assumptions about the dish surface, we take them as a representative pair of possible dish models. We use the empirically best feed height of 5.3\,m. We also include the Airy pattern for comparison as in \citet{nithya15}. Beam models with weaker response near the horizon (such as the Airy pattern) downweight sources in this direction of high apparent frequency dependence. This reduces the magnitude of emission near the edge of the EOR window, reducing the risk it leaks inside. We use the per-baseline approach of \citet{parsons12a,parsons12b} by first simulating visibilities measured by specific baselines as a function of frequency, then computing the Fourier transform over frequency (delay transform), and lastly normalizing the result into a cosmological power spectrum following \citet{nithya15}.

@@

Figure \ref{fig:delayspec} (top panel) shows simulated foreground delay spectra at various LSTs using the full-faceted beam. As all these LSTs correspond to high galactic latitudes far from the galactic center, the total visibility power (the level of the zero delay mode) varies only by a factor of a few over these LSTs on both baseline lengths (14.6\,m (left panel), 43.8\,m (right panel)). However the positive delay horizon limit (corresponding to the western horizon) has a peak that varies by over 1.5 orders of magnitude on both baselines, demonstrating the stark difference in horizon brightening when the galaxy is just above versus just below the horizon. In this figure we perform the approximate conversion from delay $\tau$ to $k_\parallel$ at $z=8$, which we plot as a second $x$-axis at the top of the plot.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{nithya_fg_pspec_all_lst.pdf}
\includegraphics[width=6in]{nithya_fg_pspec_lst0deg.pdf}
\includegraphics[width=6in]{nithya_fg_pspec_lst60deg.pdf}
\caption{We plot simulated foreground delay spectra using the full-faceted beam at various LSTs (top panel). The maximum horizon brightening at the positive horizon occurs close to 0$^\circ$ LST. At this LST, the simulated foreground delay spectra for the three beam models differ markedly near the positive horizon, plotted as a vertical line at the baseline's maximum delay. In contrast, when the horizon brightening effect is smaller at 60$^\circ$ LST (bottom panel), the foreground delay spectra from all three beams agree better.}
\label{fig:delayspec}
\end{figure*}

@@

How much do the predicted foreground power spectra differ between the three model dish power patterns? Figure \ref{fig:delayspec} (middle panel) shows the simulated delay spectra for all three beams at $0^\circ$ LST, when the horizon brightening is worst. Both numerical models agree out to delays of roughly 20\,ns on the 14.6\,m baseline and 50\,ns on the 43.8\,m baseline. These numbers suggest that the beams track each other fairly well out to roughly 25$^\circ$ from zenith, beyond which they diverge. This is roughly what is observed in Figure \ref{fig:measuredbeamslices}. At larger delays, especially near the positive delay horizon limit, all three model delay spectra diverge due to the significant edge brightening which effectively discriminates between these models. The perfect paraboloid and full-faceted beams reach roughly -32\,dB and -38\,dB at $80^\circ$ zenith angle (Figure \ref{fig:modelbeams}), consistent with the fact that the perfect paraboloid beam has a larger horizon brightening than the full-faceted beam. This is seen in the delay spectra for both baseline lengths, though the edge brightening is much clearer on the longer baseline where it less diluted by zero delay emission.

@@

We have presented beam pattern measurements at 137\,MHz, and discussed their implications for 21\,cm power spectrum analyses in terms of sensitivity and foreground isolation. We begun with power pattern measurements made using the beam mapping system of \citet{neben15} which we deployed at the prototype three-element HERA array at the National Radio Astronomy Observatory--Green Bank. We measured the dish power pattern with the feed at different heights over the dish surface and found that the best focus is at a feed rigging height of 5.3\,m, though this may change for different feed designs being explored \citep{feedoptimizationmemo}. The measured beams probe nearly two thirds of the visible sky down to -30\,dB relative to boresight, and agree well with both models in the main lobe out to 10--20$^\circ$ from zenith. The measured beams roughly track the predicted sidelobe levels at 20--30\,dB below zenith, deviating at the 1--5\,dB level.

These deviations away from models and away from 180$^\circ$ azimuthal symmetry are larger than the $\pm0.5$\,dB systematics observed in our null experiments which probe the accuracy of our beam measurement system, suggesting they are genuine measurements of the in situ dish beam. The most likely culprit is feed mis-centering which shifts and distorts the main lobe sidelobes. In the full HERA array, the suspended feeds will be tied to the dish surface at several points to fine tune the feed centering and leveling, and mitigate wind buffeting. Characterizing the level of antenna-to-antenna beam variation in the full HERA array and its effects on power spectrum analyses, as \citet{neben16} do for the MWA, is left as future work.

We quantify HERA's sensitivity to the 21\,cm power spectrum given our beam measurements by first computing the collecting area of the measured beams at the different feed rigging heights, finding 93\,m$^2$ at the best focus, implying an aperture efficiency of 60\%. Feed optimization is ongoing, but the present feed sacrifices aperture efficiency in order to taper the dipole beam towards the edges of the dish and make the X and Y dipole beams as similar as possible using a cylinder hanging from the back plane. We convert our measured collecting areas into effective dish sizes, then use 21cmSense to predict the overall power spectrum SNR at $z\sim9.5$ with one season of HERA-320 data. We predict SNRs of 12.7 and 74.3 using foreground avoidance and subtraction approaches respectively, compared with SNRs of 18.7 and 90.8 using an ideal unobstructed 14\,m aperture (Airy pattern). Still, these sensitivities permit a very significant detection of the 21\,cm signal after a single observing season.

@@

We predict the magnitude of this effect for the HERA element and discuss the uncertainties in this estimate due to beam modeling uncertainty. As expected, we find that the level of horizon brightening is largest when the galaxy is just above the horizon, and lowest when it is well below. When this pitchfork effect is large, we find that the uncertainty in its predicted amplitude is also large, as seen in the differences between the delay spectra calculated using full-faceted and perfect paraboloid beam models. When the effect is small, the two beam models produce much more similar results, highlighting the delay spectrum as an exquisite probe of the difficult-to-measure beam response at very low elevations. Of course the delay spectrum provides only an integrated measure of the beam, but some information can still be extracted. By forward modeling foreground delay spectra using different MWA primary beam models, for instance, it was observed that the MWA bowtie dipoles are better modeled as isotropic radiators than hertzian dipoles at these low elevations (N. Thyagarajan, private communication). Direct measurements using transmitter-equipped drones would be ideal and their development is ongoing \citep{drone1,drone2}.
